Computing Entropies With Nested Sampling

### 2017-07-12 paper/ms.tex
Sign fix
Fixed errors in Poisson process part

To overcome this, NS evolves a population of $N$ particles in the parameter
space towards higher likelihood regions.
The particles are initially initialised from the prior $\pi(\x)$, and the
particle with the lowest likelihood, $L^*_1$, is found and its details are
recorded as output. This worst particle is then discarded
and replaced by a new particle
drawn from the prior $\pi$ but subject to the constraint that its likelihood
must be above $L^*_1$ (this is usually achieved by cloning a surviving particle
and evolving it with MCMC to explore
$\pi(\x)\mathds{1}\left(L(\x) > L^*_1\right)$).
This process is repeated, resulting in an
increasing sequence of likelihoods
\begin{align}
L^*_1, L^*_2, L^*_3,... \label{eqn:likelihood_sequence}
\end{align}
Defining $X(\ell)$ as the amount of prior mass with likelihood greater than
some threshold $\ell$,
\begin{align}
X(\ell) &= \int \pi(\x) \mathds{1}\left(L(\x) > \ell\right) \, d\x,
\end{align}
each discarded point in the sequence can be assigned an $X$-value,
which transforms $Z$ to a one-dimensional integral.
\citet{skilling2006nested}'s key insight was that the
$X$-values of the discarded points are approximately known from the nature of the
algorithm. Specifically, each iteration reduces the prior mass by approximately
a factor $e^{-1/N}$. More accurately, the conditional distribution of
$X_{i}$ given $X_{i-1}$ (and the definition of the algorithm)
is obtainable by letting
$t_i \sim \textnormal{Beta}(N,1)$ and setting $X_{i+1} := t_iX_{i-1}$.
The $t_i$ variables represent the proportion of the remaining prior mass
that is {\em retained} at each iteration, after imposing the constraint of the
latest $L^*$ value.

@@

Consider $1-t_i$, the fraction of the remaining prior mass that is {\em removed}
in each NS iteration.
The distribution for the compression factors $t_i \sim $Beta$(N,1)$
corresponds to an exponential distribution for $\ln(1-t_i)$:
\begin{align}
\ln (1 - t_i) &\sim \textnormal{Exponential}(N).
\end{align}

Therefore,
the sequence of $-\ln X$ values produced by a Nested Sampling run can be
considered as a Poisson process with rate equal to $N$, the number of
particles. This is why separate NS runs can be simply merged
\citep{skilling2006nested}.
\citet{Walter2015} showed how this view of the NS sequence of points
can be used to construct a
version of NS that produces unbiased (in the frequentist sense)
estimates of the evidence.
However, it can also be used to construct an unbiased estimator of
a log-probability, which is more relevant to information theoretic
quantities discussed in the present paper.

>>>>

To overcome this, NS evolves a population of $N$ particles in the parameter
space towards higher likelihood regions.
The particles are initially initialised from the prior $\pi(\x)$, and the
particle with the lowest likelihood, $L^*_1$, is found and its details are
recorded as output. This worst particle is then discarded
and replaced by a new particle
drawn from the prior $\pi$ but subject to the constraint that its likelihood
must be above $L^*_1$ (this is usually achieved by cloning a surviving particle
and evolving it with MCMC to explore
$\pi(\x)\mathds{1}\left(L(\x) > L^*_1\right)$).
This process is repeated, resulting in an
increasing sequence of likelihoods
\begin{align}
L^*_1, L^*_2, L^*_3,... \label{eqn:likelihood_sequence}
\end{align}
Defining $X(\ell)$ as the amount of prior mass with likelihood greater than
some threshold $\ell$,
\begin{align}
X(\ell) &= \int \pi(\x) \mathds{1}\left(L(\x) > \ell\right) \, d\x,
\end{align}
each discarded point in the sequence can be assigned an $X$-value,
which transforms $Z$ to a one-dimensional integral.
\citet{skilling2006nested}'s key insight was that the
$X$-values of the discarded points are approximately known from the nature of the
algorithm. Specifically, each iteration reduces the prior mass by approximately
a factor $e^{-1/N}$. More accurately, the conditional distribution of
$X_{i}$ given $X_{i-1}$ (and the definition of the algorithm)
is obtainable by letting
$t_i \sim \textnormal{Beta}(N,1)$ and setting $X_{i+1} := t_iX_{i-1}$ (where
$X_1 = t_1$.
The $t_i$ variables represent the proportion of the remaining prior mass
that is {\em retained} at each iteration, after imposing the constraint of the
latest $L^*$ value.

@@

The sequence of $X$ values is defined by
\begin{align}
X_1 &= t_1 \\
X_2 &= t_2X_1 \\
X_3 &= t_3X_2
\end{align}
and so on, i.e. $X_i$ can be written as
\begin{align}
X_i = \prod_{k=1}^i t_k.
\end{align}
Taking the log and multiplying both sides by -1, we get
\begin{align}
-\ln X_i = \sum_{k=1}^i \left(-\ln t_k\right).\label{eqn:minuslogX}
\end{align}

Consider $t_i$, the fraction of the remaining prior mass that is retained
in each NS iteration. By a change of variables,
the distribution for these compression factors $t_i \sim $Beta$(N,1)$
corresponds to an exponential distribution for $-\ln(t_i)$:
\begin{align}
-\ln t_i &\sim \textnormal{Exponential}(N).\label{eqn:exp}
\end{align}

From Equations~\ref{eqn:minuslogX} and~\ref{eqn:exp}, we see that
the sequence of $-\ln X$ values produced by a Nested Sampling run can be
considered as a Poisson process with rate equal to $N$, the number of
particles. This is why separate NS runs can be simply merged
\citep{skilling2006nested}.
\citet{Walter2015} showed how this view of the NS sequence of points
can be used to construct a
version of NS that produces unbiased (in the frequentist sense)
estimates of the evidence.
However, it can also be used to construct an unbiased estimator of
a log-probability, which is more relevant to information theoretic
quantities discussed in the present paper.
